From Light to Growth: A Study of Night-Time Light as an Indicator of Economic Development

### 2024-12-02 paper/paper.qmd
Wrote the modeling section.

We collected data from multiple sources to analyze the relationship between night-time light intensity and GDP growth. The primary dataset is a harmonized global night-time light (NTL) dataset spanning from 1992 to 2018, developed by Li et al. [@Li2020] The dataset combines observations from two satellite systems: DMSP/OLS (1992–2013) and VIIRS (2012–2018). To ensure consistency across years, we addressed differences in spatial resolution and saturation between the two systems by inter-calibrating DMSP data and simulating DMSP-like VIIRS observations. The resulting dataset provides annual Digital Number (DN) values, representing night-time illumination intensity, as GeoTIFF files.

We also sourced GDP, manufacturing share of GDP, and population data from the World Bank’s World Development Indicators (WDI) [@WorldBankWDI],measuring GDP in purchasing power parity (PPP) for comparability across countries. Additionally, we incorporated the Statistical Performance Indicator (SPI), which evaluates national statistical systems across five pillars: data use, data services, data products, data sources, and data infrastructure. [@WorldBankSPI]

@@

The DN (Digital Number) values in the dataset are derived from satellite imagery that captures night-time light emissions. These emissions result from real-world activities such as streetlights, industrial complexes, and residential areas. The satellite sensors detect these light intensities, which are then processed, calibrated, and harmonized to produce standardized measurements. These measurements are ultimately converted into numerical entries representing the brightness of light observed in specific geographical locations, enabling researchers to analyze economic and social activity at various scales. The data set for our analysis includes NTL captured by the Defense Meteorological Satellite Program’s Operational Linescan System (DMSP/OLS) and the Visible Infrared Imaging Radiometer Suite (VIIRS), provides a spatially explicit proxy for human presence. DMSP/OLS, operational since the 1970s, used low-light imaging to detect artificial lights but faced challenges like coarse resolution and sensor saturation [@elvidge1997]. VIIRS, introduced in 2012, improved spatial resolution and dynamic range, using advanced algorithms to remove noise and non-anthropogenic light sources, resulting in high-quality radiance-calibrated maps [@elvidge2017].

The World Bank datasets, such as GDP, population, and manufacturing share of GDP, are sourced from national statistical agencies and harmonized using internationally accepted methodologies like the System of National Accounts (SNA). GDP is measured through national accounts using production, income, and expenditure approaches, relying on administrative records, business surveys, and sectoral data. Population data originate from national censuses and are updated between censuses using demographic surveys, vital registration systems, and migration records. Data on the manufacturing share of GDP are collected through industrial surveys and administrative records, capturing outputs from manufacturing activities.

## Data Extraction and Cleaning

@@

To associate each observation with a specific country, the extracted data were merged with a global shapefile of country boundaries through spatial join techniques. Each latitude-longitude pair was assigned to the country in which it was located, and observations without an assigned country—primarily from international waters—were removed. Subsequently, to create an aggregate annual measure of NTL intensity for each country, all observations within a country were summed to calculate the total light emitted annually.

@@

library(arrow)        
library(tidyverse)         
library(modelsummary)
library(here)
library(kableExtra)
library(tinytable)
library(sf)
library(scales)
library(ggridges)
library(gridExtra)
library(gridExtra)
library(grid)
library(png)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
# Load the shapefile and suppress messages
shapefile_path <- "/Users/shamayladurrin/Desktop/STA304/NTLGDPCorr/data/01-raw_data/02-shapefile"
sink("/dev/null")  # Redirect console output to null
world_shapefile <- st_read(shapefile_path)
sink()# Load the Parquet file
metrics <- read_parquet(here("data/03-plotting_data/plotting.parquet"))
```

@@

knitr::include_graphics("/Users/shamayladurrin/Desktop/STA304/NTLGDPCorr/others/plots/test.png")

@@

@@

>>>>

We collected data from multiple sources to analyze the relationship between night-time light intensity and GDP growth. The primary dataset is a harmonized global night-time light (NTL) dataset spanning from 1992 to 2018, developed by Li et al. [@Li2020] The dataset combines observations from two satellite systems: DMSP/OLS (1992–2013) and VIIRS (2012–2018). To ensure consistency across years, they addressed differences in spatial resolution and saturation between the two systems by inter-calibrating DMSP data and simulating DMSP-like VIIRS observations. The resulting dataset provides annual Digital Number (DN) values, representing night-time illumination intensity, as GeoTIFF files.

We also sourced GDP, manufacturing share of GDP, and population data from the World Bank’s World Development Indicators (WDI) [@WorldBankWDI],measuring GDP in purchasing power parity (PPP) for comparability across countries. Additionally, we incorporated the Statistical Performance Indicator (SPI), which evaluates national statistical systems across five pillars: data use, data services, data products, data sources, and data infrastructure. [@WorldBankSPI] A thorough discussion on how we organised the data set is given in [@sec-clean].

@@

Night-time light intensity, which reflects human activity, is indirectly measured by satellite sensors onboard the DMSP/OLS and VIIRS systems. These satellites capture light emissions from streetlights, industrial complexes, and residential areas, converting this physical phenomenon into Digital Number (DN) values. The brightness levels are then transformed into quantifiable data, which are stored in GeoTIFF format for further analysis. DMSP/OLS, operational since the 1970s, used low-light imaging to detect artificial lights but had limitations such as coarse resolution and sensor saturation [@elvidge1997]. VIIRS, launched in 2012, provided better spatial resolution and could capture a wider range of light levels, with improved algorithms to filter out unwanted noise and natural light sources.[@elvidge2017]. The harmonized version of these two datasets, combining data from DMSP/OLS and VIIRS, has been used by researchers for analysis.

GDP is measured as the total economic output of a country, adjusted for purchasing power parity (PPP) to standardize across different national economies. This transformation from an elaborate economic reality into a single numerical value is done through extensive national accounting systems, with data collected by governments through economic surveys, reports, and censuses.Population figures, reflecting demographic changes and the size of human capital, are also derived from national censuses and are measured in absolute numbers. Meanwhile, the manufacturing share of GDP is calculated by extracting relevant economic data, representing the portion of a country's GDP generated by the manufacturing sector. The World Bank sources all these data from national accounts.

## Data Extraction and Cleaning {#sec-clean}

@@

To associate each observation with a specific country, the extracted data were merged with a global shapefile of country boundaries through spatial join using [@geopandas]. Each latitude-longitude pair was assigned to the country in which it was located, and observations without an assigned country—primarily from international waters—were removed. Subsequently, to create an aggregate annual measure of NTL intensity for each country, all observations within a country were summed to calculate the total light emitted annually.

@@

library(arrow)        
library(tidyverse)         
library(modelsummary)
library(here)
library(kableExtra)
library(tinytable)
library(sf)
library(scales)
library(gridExtra)
library(grid)
library(png)
library(modelsummary)
library(broom)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
# Load the shapefile and suppress messages
shapefile_path <- "data/01-raw_data/02-shapefile"
sink("/dev/null")  # Redirect console output to null
world_shapefile <- st_read(here(shapefile_path))
sink()# Load the Parquet file
metrics <- read_parquet(here("data/03-plotting_data/plotting.parquet"))
```

@@

knitr::include_graphics(here("others/plots/ntl.png"))

@@

# Linear Models

In this section, we present four models to assess the relationship between Night-Time Light (NTL) intensity and various economic indicators, with a particular focus on understanding how NTL correlates with GDP and how this relationship varies across countries with differing levels of statistical capacity. A key challenge in our analysis was the high correlation between GDP, manufacturing share of GDP, and population, as highlighted in the correlation matrix provided in the appendix. This multicollinearity made it impractical to include these variables in the same model without risking inflated standard errors and unreliable coefficients. To address this, we developed separate models for each variable.

The first model examines the relationship between NTL intensity and GDP, incorporating country-specific and year-specific fixed effects. The inclusion of country-specific fixed effects accounts for unobserved factors that differ across countries, such as geographic characteristics, governance structures, or cultural differences. Year-specific fixed effects are included to capture variations over time, such as global economic trends or technological advancements, that may affect the relationship between NTL and GDP.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(GDP_{i,t}) + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$ Country fixed effects
  - $gamma_t$ Year fixed effects

Building on the first model, the second model investigates how the relationship between NTL and GDP varies across countries with differing statistical capacities. As seen in Figure 1, the scatterplot of GDP and NTL faceted by statistical capacity suggests a weaker correlation in countries with lower statistical capacity. To explore this further, we include an interaction term between GDP and statistical capacity grade, allowing us to examine whether countries with stronger statistical systems exhibit a more consistent and robust relationship between NTL and GDP. The intuition is that higher statistical capacity is associated with better data quality and more stable economic structures, which may enhance the reliability of NTL as a proxy for economic activity. Similar to the first model, we control for year-specific and country-specific fixed effects to account for temporal and cross-country heterogeneity.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(GDP_{i,t}) + \beta_2 (\text{Grade}_{i}) + \beta_3 (\log(GDP_{i,t}) \times \text{Grade}_{i}) + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - Interaction term: $$ \log(\text{GDP}) \times \text{Grade} $$
  - $gamma_t$: Year fixed effects

The third model examines the relationship between NTL intensity and population, while controlling for country and year fixed effects. Population is a key driver of light emissions, particularly in urban areas, and this model isolates its unique contribution to NTL by excluding GDP to address multicollinearity. This approach helps assess whether NTL can serve as a proxy for population, especially in regions where population data may be less reliable, while still accounting for variations across countries and years.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(Population_{i,t}) + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$: Country fixed effects
  - $gamma_t$: Year fixed effects

The fourth model examines the relationship between NTL intensity and the manufacturing share of GDP, controlling for country and year fixed effects. Manufacturing activity, often concentrated in industrial areas, significantly influences light emissions. By isolating this variable, the model highlights how industrialization shapes NTL patterns, independent of overall GDP, while accounting for country- and time-specific factors.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \text{Manufacturing Share}_{i,t} + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$: Country fixed effects
  - $gamma_t$: Year fixed effects

# Model Summary

#| echo: false
#| warning: false
#| message: false
#| label: tbl-summodel
#| fig-cap: "Model Summary"

model1 <- readRDS(here("models/model1.rds"))
model2 <- readRDS(here("models/model2.rds"))
model3 <- readRDS(here("models/model3.rds"))
model4 <- readRDS(here("models/model4.rds"))
# List of your models
models <- list(
  model1,
  model2,
  model3,
  model4
)

# Function to clean model output by removing fixed effects
clean_model <- function(model) {
  tidy(model) %>%
    filter(!grepl("factor\\(country\\)|factor\\(year\\)", term)) %>% # Remove fixed effects coefficients
    rename(
      Term = term,
      Estimate = estimate,
      `Std. Error` = std.error,
      `t-value` = statistic,
      `p-value` = p.value
    )
}

# Generate a cleaned version of the model summaries
cleaned_models <- lapply(models, clean_model)

# Create a modelsummary table, keeping AIC and BIC
model_summary_table <- modelsummary(
  models,
  coef_omit = "factor\\(country\\)|factor\\(year\\)", # Exclude fixed effects
  gof_map = c("r.squared", "adj.r.squared", "AIC", "BIC", "nobs") # Include key statistics
)

# Display the table
model_summary_table
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: scatterplot-matrix
#| fig-cap: "Scatterplot matrix showing pairwise relationships between key variables in the analysis_data dataset."

# Load necessary library
library(GGally)
library(ggplot2)

analysis_transformed <- analysis_data %>%
  mutate(
    log_dn = log(dn),
    log_gdp = log(gdp),
    log_population = log(population)
  ) %>%
  filter(complete.cases(.)) %>%
  select(log_dn, log_gdp, manufacturingsharegdp, log_population)

# Create the scatterplot matrix
ggpairs(
  analysis_transformed,
  title = "Scatterplot Matrix of Key Variables",
  lower = list(
    continuous = wrap("points", alpha = 0.4, size = 0.1) # Smaller, semi-transparent points
  ),
  diag = list(
    continuous = wrap("barDiag", color = "white", fill = "gray50", alpha = 0.8) # Cleaner histograms
  ),
  upper = list(
    continuous = wrap("cor", size = 4, alignPercent = 0.8) # Correlation coefficients
  )
) +
  theme_minimal(base_size = 10) + # Apply a minimal theme
  theme(
    panel.grid = element_blank(), # Remove gridlines for a cleaner look
    axis.text = element_text(size = 8), # Adjust axis text size
    axis.title = element_text(size = 10, face = "bold") # Bold axis titles
  )
